---
title: "Network analysis demo"
---

Networks (also called graphs) are a way of studying connections. Many different kinds of things that we might want to study historically can be connections. For instance, correspondence between individuals can show the structure of a group. Migration between countries is a kind of a network. So too are borrowings between texts. Once you understand the concept of connections and the typical structure of network data, you will find networks everywhere.

In addition to the tidyverse package, we will need three other packages. The igraph package provides the low-level understanding of networks and how to analyze then. The tidygraph package lets us manipulate network data using the tidyverse style functions we are used to. And the ggraph package lets us plot networks using ggplot2. You can install these with `install.packages(c("tidygraph", "ggraph", "igraph"))`.

We start by loading the packages.

```{r setup, include=FALSE}
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
```

We are going to work with a network graph that shows the level of state to state borrowings of codes of civil procedure in the nineteenth century. For example, California heavily borrowed its laws from New York, and Oregon borrowed its laws from California.

## Loading the data and creating the graph

Networks are defined by a data frame of edges and (optionally) a data frame of nodes. An edge table has two columns of node names showing a connection between them. A node table has one column of node names. Optionally, both can include additional columns of metadata specific to nodes or edges.

```{r, message=FALSE}
edges_df <-read_csv("field-states-edges.csv")
nodes_df <-read_csv("field-states-nodes.csv")
```

We can turn those data frames into a graph (the mathematical term for a network).

```{r}
g <- graph_from_data_frame(edges_df, directed = TRUE, vertices = nodes_df)
g <- g %>% as_tbl_graph()
g
```

We can get a quick look at the graph using `plot()`

```{r}
plot(g)
```

## Measuring the centrality of nodes

The importance, called *centrality*, of a node can be measured with various algorithms. These can be added to the graph object as node attributes using `mutate()`. Note that below we pull out the nodes data frame to inspect it and do not save the results back to our graph object `g` just yet.

Look at the definitions given in *The Historian's Macroscope* and follow the links to the documentation for the various functions using `?centrality`. Also use your intution from the visualization of the graph above. What measure of centrality is most appropriate for this graph?

```{r}
centrality <- g %>% 
  mutate(
    c_degree = centrality_degree() %>% round(4),
    c_betweenness = centrality_betweenness() %>% round(4),
    c_closeness = centrality_closeness() %>% round(4),
    c_pagerank = centrality_pagerank() %>% round(4),
    # c_eigen = centrality_eigen() %>% round(4),
    # c_hub = centrality_hub() %>% round(4),
    # c_communicability = centrality_communicability() %>% round(4)
         ) %>% 
  as_data_frame(what = "vertices") 
centrality
```

## Plotting

Networks can be plotted within the ggplot2 framework using the ggraph package. You will need to include geoms for both edges (`geom_edge_*`) and nodes (`geom_node_*`)

```{r}
ggraph(g, layout = "nicely") +
  geom_edge_fan(arrow = arrow(ends = "first")) +
  geom_node_point(aes(color = region), size = 10) +
  geom_node_text(aes(label = name)) +
  theme_graph()
```

## Clustering

```{r}
groups <- g %>% 
  mutate(infomap = as.factor(group_infomap()),
         walktrap = as.factor(group_walktrap()),
         components = as.factor(group_components()),
         betweenness = as.factor(group_edge_betweenness()))

ggraph(groups, layout = "nicely") +
  geom_edge_fan() +
  geom_node_point(aes(color = components), size = 10) +
  geom_node_text(aes(label = name)) +
  theme_graph()
```

## Node-pair level measures

```{r}
distances <- g %>% 
  mutate(is_ny = name == "NY",
         distance = node_distance_from(is_ny, mode = "in")) 

ggraph(distances, layout = "nicely") +
  geom_edge_fan(arrow = arrow(ends = "first")) +
  geom_node_point(aes(color = as.factor(distance)), size = 10) +
  geom_node_text(aes(label = name)) +
  theme_graph()
```

## Component graphs

Sometimes parts of a graph are unconnected to other graphs, and it would be nice to pull them out into separate pieces.

```{r}
components <- g %>% to_components()

plot(components[[1]])
plot(components[[2]])
plot(components[[3]])
```

## Bipartite graphs

A bipartite graph is a graph where the nodes fall into two different groups, and where members of a group are never connected to one another. A common example is a membership graph, with organizations and members. We can load a membership graph for revolutionary organizations in Boston.

```{r, message=FALSE}
revere_edges <-read_csv("paul-revere-edges.csv")
revere_bip <- graph_from_data_frame(revere_edges)
```

Now that we have the graph loaded we can project it. We have to distinguish between the `type` of the nodes---which belong to which side of the projection.

```{r}
V(revere_bip)$type <- bipartite_mapping(revere_bip)$type
table(V(revere_bip)$type)
bipartite_projection_size(revere_bip)
```
In this case `TRUE` will correspond to an organization and `FALSE` to an individual. We should also check how big our resulting graphs would be.

```{r}
revere_org <- bipartite_projection(revere_bip, which = "true")
plot(revere_org)
```

## In-class assignment

Option 1: Redo some of these analyses using the Field Code code-to-code level data. Try plotting the codes by a centrality measure. Get as close as you can to a presentation graphic.

- `field-codes-edges.csv`
- `field-codes-nodes.csv`

Option 2: Reproduce Kieran Healy's analysis from this [blog post](https://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/). Use `paul-revere-edges.csv`. Get as close as you can to a presentation graphic.
