---
title: "Finding Paul Revere"
output: html_document
---

The goal of this assignment is to work with a list of members of various revolutionary organizations in Boston and do some network analysis on it. Disclaimer: the idea for this assignment and the data are [borrowed from Kieran Healy](https://github.com/kjhealy/revere), but don't look at that link until the assignment is complete. 

```{r setup, include=FALSE}
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
```

## Load the data and create a network graph

This loads the edges data frame.

```{r, message=FALSE}
edges_df <- read_csv("paul-revere-edges.csv")
```

Now turn the edges of people-to-organization into a graph object, and use the sample code to project it into a graph of people-to-people connections. Remember that to use the tidygraph functions you need to run `as_tbl_graph` on the graph object. 

```{r}

```

## Project the bipartite network graph 

The graph that you have is bipartite---it shows people's relationships to organizations. Project it into a person-to-person network. Use the sample code from the demo file.

```{r}

```

## Measure the centrality of the nodes

Measure the centrality of the nodes (i.e., the people). You will have to try several centrality measures and pick one that is most representative to the task. Degree centrality (i.e., the number of edges attached to a node) is the simplest measure of centrality.

```{r}

```

Which centrality measure did you find most reliable and why?

Whose names turned up that you recognized? That you did not recognize? What would you do next?

## Create a plot of the network

Plot the network graph using the ggraph package and the sample code from the demo. (Youn can get a quick and dirty look at the network using `plot(my_graph, vertex.label = NA)`.) Remember that you will need to add a `geom_node_*` and `geom_edge_*` to plot both of those.

```{r}

```

The full graph is almost certainly too difficult to read initially. Try changing the `alpha` (transparency) on the edges as a first step. 

Another approach is to present a simplified version of the network graph with just the most important people. One way that you could do this is by filtering out people who have a stronger connection. The `weight` attribute of the edges of the graph is a rough estimate of this property. You can use dplyr verbs like `filter()` on the graph object if you first tell it whether you want to apply the function to the nodes or the edges: `my_graph %>% activate(edges) %>% filter(weight > 1)`. Note that there is a function `node_is_isolated()` which will tell you if a node is completely disconnected.

```{r}

```
